---
title: "SM_Processing_Dec2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

source("SM_Processing_Functions.R")

```

############# RCM compile for July 2021 completed by Simon on 7/8/2021 because I need this data for my thesis, Childs, Control, Marian still need compilation #########################

#### Rock Creek Soil Moisture Data

#### Prev. Rock Creek Compile (September 2021)

```{r}

RC_compile_Sept2021 <- readr::read_csv(file = "RC_SM_Sept2021/compiled_RC_Sept2021.csv", 
                                       col_types = cols(Date = col_character(), .default = col_double())) %>% 
  dplyr::mutate(Date = lubridate::ymd_hms(Date))

```

#### New Compile for RC (December 2021)

```{r}

compiled_RC_Dec2021 <- process_SM_data(path = "RC_SM_Dec2021", 
                                        site = "RC", prev_compile = RC_compile_Sept2021, written_file_name = "compiled_RC_Dec2021.csv")
```
#### Data Cleaning Extra Step (December 2021)

```{r}

RC_compile_ver2_Dec2021 <- readr::read_csv(file = "RC_SM_Dec2021/compiled_RC_Dec2021.csv", 
                                       col_types = cols(Date = col_character(), .default = col_double()))%>%
      dplyr::distinct()%>%

readr::write_csv(., path = paste0("RC_SM_Dec2021", "/", written_file_name = "compiled_ver2_RC_Dec2021.csv"))
```

#### Prev. Compile Marian (Sept 2021)

```{r}
MM_compile_Sept2021 <- readr::read_csv(file = "Marian_SM_Sept2021/compiled_Marian_Sept2021.csv", 
                                       col_types = cols(Date = col_character(), .default = col_double())) %>% 
  dplyr::mutate(Date = lubridate::ymd_hms(Date))
```

#### New Compile for Marian (December 2021)

```{r}
compiled_Marian_Dec2021 <- process_SM_data(path = "Marian_SM_Dec2021", 
                                        site = "marian", prev_compile = MM_compile_Dec2021, written_file_name = "compiled_Marian_Dec2021.csv")
```


#### Prev. Compile Control (September 2021)

```{r}
Control_compile_Sept2021 <- readr::read_csv(file = "Control_SM_Sept2021/compiled_Control_Sept2021.csv", 
                                       col_types = cols(Date = col_character(), .default = col_double())) %>% 
  dplyr::mutate(Date = lubridate::ymd_hms(Date))
```

#### New Compile for Control (December 2021)

```{r}
compiled_Control_Dec2021 <- process_SM_data(path = "Control_SM_Dec2021", 
                                        site = "control", prev_compile = Control_compile_Sept2021, written_file_name = "compiled_Control_Dec2021.csv")
```

#### Prev. Compile Child's (May 2021)

```{r}
Childs_compile_May2021 <- readr::read_csv(file = "Childs_SM_May2021/compiled_Childs_May2021.csv", 
                                       col_types = cols(Date = col_character(), .default = col_double())) %>% 
  dplyr::mutate(Date = lubridate::ymd_hms(Date))
```

#### New Compile for Child's (July 2021)

```{r}
compiled_Childs_July2021 <- process_SM_data(path = "Childs_SM_July2021", 
                                        site = "Childs", prev_compile = Childs_compile_May2021, written_file_name = "compiled_Childs_July2021.csv")
```
